---
title: "Article 1: Digital Trace Data as Indicators of Social Data Validating Google Trends for  use in Scientific Research"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
output:
  pdf_document: 
    keep_tex: true
    latex_engine: xelatex
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{longtable}
   - \usepackage{caption}
   - \usepackage{booktabs}
   - \usepackage{placeins}
   - \usepackage{hhline}
   - \doublespacing
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      cache = TRUE,
                      fig.width = 7,
                      fig.asp = 0.8,
                      out.width = "80%",
                      fig.align="center"
                      )

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales, nlme,
               flextable)


my_palette = c('#56b4e9','#999999','#E69F00','#009E73','#CC79A7','#0072B2','#D55E00','#E69F00')
my_palette  <- c(MetBrewer::met.brewer(name = 'Cross', type = 'discrete'), '#555F61')

## Automatically use showtext to render text
font_add_google("Lora", "lora")
showtext_auto()

theme_diss <- function(base_size = 14) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      # Figure assembly
      plot.title = element_text(family="lora",
                                size = rel(1), 
                                margin = margin(0,0,5,0), 
                                hjust = 0),
      plot.title.position = "plot",
      plot.subtitle = element_text(family="lora",
                                   size = rel(0.85)),
      plot.caption = element_text(family="lora",
                                  size = rel(0.70),
                                  hjust = 1),
      # Graphical Zones
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Axes
      axis.title = element_text(family="lora", 
                                size = rel(0.85)),
      axis.text = element_text(family="lora", 
                               size = rel(0.70)),
      axis.line = element_line(color = "black", 
                               arrow = arrow(length = unit(0.3, "lines"), 
                                             type = "closed")),
      # Legend
      legend.title = element_text(family="lora",
                                  size = rel(0.85)),
      legend.text = element_text(family="lora", 
                                 size = rel(0.70)),
      legend.key = element_rect(fill = "transparent", 
                                colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", 
                                       colour = NA),
      # Facetting 
      strip.background = element_rect(fill = "#17252D",
                                      color = "#17252D"),
      strip.text = element_text(family="lora",
                                size = rel(0.85), 
                                color = "white", 
                                margin = margin(5,0,5,0))
    )
}

theme_set(theme_diss()) 

```
```{r data}
#| cache = FALSE


vacc_hes <- read_rds(here('data', 'cleaned_data', 'vacc_hes.rds'))
mask <- read_rds(here('data', 'cleaned_data', 'mask.rds')) %>%  ungroup()
covid <- read_rds(here('data', 'cleaned_data', 'covid.rds'))
suicide <- read_rds(here('data', 'cleaned_data', 'suicide.rds'))
pres_2016 <- read_rds(here('data', 'cleaned_data', 'pres_2016.rds'))
pres_2020 <- read_rds(here('data', 'cleaned_data', 'pres_2020.rds'))
covid_corrs <- read_rds(here('data', 'from_hpc', 'covid_corrs.rds'))
suicide_corrs <- read_rds(here('data', 'from_hpc', 'suicide_corrs.rds'))


```

# Research Methodology

```{r data-sources-table}
#| results='asis'

data_sources <- tibble::tribble(
~type, ~Validated.Data.Source, ~Type, ~Dates, ~Google.Trends.Used,
"Behaviors and Attitudes","General Social Survey", "Cross-Sectional", "2010 - 2020", NA,
 
"Behaviors and Attitudes","Vaccine Hesitancy for COVID-19", "Cross-Sectional", "March 3 – 15, 2021", "Search Topics: 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Vaccine'",
 
"Behaviors and Attitudes","Mask-Wearing Survey Data", "Cross-Sectional", "July 2 - 14, 2020", "Search Topics: 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth Face Mask', 'Mask', 'Civil and Political Rights'",
 
"Health","Covid Rates", "Longitudinal", "Every Monday, 2020 - 2021", "'Covid-19', 'Coronavirus', 'Taste Loss', 'Smell Loss'",
 
"Health","County Suicide Rates", "Longitudinal", "Yearly 2010-2020", "Search Topics: 'Suicide', 'Depression', Search Term:' Suicide Hotline'",

'Political',"American National Election Survey", "Cross-Sectional", "2020", NA,
 
'Political',"Presidential Election Results", "Cross Sectional", "2016 & 2020", "Search Topics: 'Hilary Clinton', 'Donald Trump', 'Joe Biden'"
 ) %>% as_grouped_data(groups = c("type"), columns = NULL)


data_sources %>% 
  as_flextable(hide_grouplabel = TRUE) %>% 
  set_caption(caption = "New York Air Quality Measurements")  %>% 
  # theming
  theme_box() %>% 
  bg(bg = my_palette[8], part = "header") %>% 
  color(color = "white", part = "header") %>%
  set_table_properties(layout = "autofit", width = 1) %>% 
  # grouped row style
  fontsize(size = 8) %>% 
  bold(i = ~ !is.na(type), bold = TRUE) %>% 
  italic(i = ~ !is.na(type), italic = TRUE) %>% 
  bg(i = ~ !is.na(type), bg = "#bad7db", part = "body")

```

## Measures

```{r}
desc_stats_table <- 
  vacc_hes  %>% 
  bind_rows(mask) %>% 
  bind_rows(covid) %>% 
  bind_rows(suicide) %>% 
  bind_rows(pres_2016 ) %>% 
  bind_rows(pres_2020 ) 

datasummary(formula = 
          (`Rate of Vaccine Hesitancy` = vacc_hesitant) +             
          (`Rate of Rare Mask Usage` = mask_rare) +   
          (`Rate of Covid` = covid_rate) +                 
          (`Rate of Deaths by Suicide` = death_rate) +                
          (`Percent of votes for Joe Biden` = biden_p) +                   
          (`Percent of votes for Donald Trump` = trump_p) +                    
          (`Percent of votes for Hillary Clinton` = clinton_p) + 
          (`Search for Covid-19` = covid_19) +                  
          (`Search for Covid-19 Vaccine` = covid_19_vaccine) +
          (`Search for Vaccine` = vaccine) +
          (`Search for coronavirus virus` = coronavirus_virus) +
          (`Search for coronavirus disease` = coronavirus_disease) +
          (`Search for Civil and political rights` = civil_and_political_rights) +
          (`Search for cloth face mask` = cloth_face_mask) +
          (`Search for Mask` = mask) +
          (`Search for Taste Loss` = taste_loss) +                 
          (`Search for Smell Loss` = smell_loss) + 
          (`Search for Suicide` = suicide) +                    
          (`Search for Depression` = depression) +                
          (`Search for Suicide Hotline` = suicide_hotline) +  
          (`Search for Hillary Clinton` = hillary_clinton_trend) +     
          (`Search for Donald Trump` = donald_trump_trend) +        
          (`Search for Joe Biden` = joe_biden_trend) +         
          (`Total Population` = total_pop) +  
          (`Population Density` = pop_density) +                
          (`Unemployment Rate` = unemployment_rate) +         
          (`% over 65` = over_65) +                   
          (`% below poverty line` = poverty_rate) +               
          (`Median income` = median_income) +             
          (`% with broadband` = broadband)  ~ 
    (Min  + Max + Mean + SD) ,
  data = desc_stats_table,
  fmt = '%.2f',
  title = 'Descriptive Statistics for Dichotomous and Numeric Variables',
  output = "latex")

```


#### vacc_hes

`r nrow(vacc_hes)`

from `r percent_format(0.01)(min(vacc_hes$vacc_hesitant))` to 
`r percent_format(0.01)(max(vacc_hes$vacc_hesitant))`.  

#### mask
(range = `r percent_format(0.01)(min(mask$mask_rare))` to  
`r percent_format(0.01)(max(mask$mask_rare))`)

### health
#### covid
0 to a maximum of `r max(covid$covid_rate)` for each 
Monday from `r glue("{format( min(covid$date), '%B %d, %Y')} through {format(max(covid$date), '%B %d, %Y')}")`.
There are `r nrow(covid)` cases across `r nrow(count(covid, fips))` 
counties and `r nrow(count(covid, date))` dates. 

#### suicide
 There are `r nrow(suicide)` total
cases, resulting from `r nrow(count(suicide, year))` observations of 
`r nrow(count(suicide, fips))` counties. Missing data were interpolated using.

Measures range from `r round(min(suicide$death_rate, na.rm=T),3)` to 
`r round(max(suicide$death_rate, na.rm=T),3)`.

### political
 avaiable for `r nrow(pres_2016)` counties in 2016   
 and `r nrow(pres_2020)` counties in 2020. 
 
with the lowest percent at 
`r percent_format(0.01)(min(pres_2020$biden_p, na.rm=T))`
and the highest at `r percent_format(0.01)(max(pres_2020$biden_p, na.rm=T))`. 



```{r corr-results}
#|  results='asis'

options(knitr.kable.NA = '')

table2 <- tribble(
  ~'grp', ~'variable', ~'measure',~'trend1', ~'trend2', ~'trend3', ~'trend4', ~'trend5', 
  "Vaccine Hesitancy", NA,  NA, 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Vaccine', NA,
  "Vaccine Hesitancy", "Vaccine Hesitancy", "Pearson's R Correlation",
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$covid_19_vaccine, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$coronavirus_disease, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$coronavirus_virus, use = "complete.obs"),4)),
  as.character(round(cor(vacc_hes$vacc_hesitant, vacc_hes$vaccine, use = "complete.obs"),4)), NA,
  
  "Mask Attitudes",NA,  NA, 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth Face Mask', 'Mask', 'Civil and Political Rights',
  "Mask Attitudes","Mask Rare", "Pearson's R Correlation",
  as.character(round(cor(mask$mask_rare, mask$coronavirus_disease, use = "complete.obs"),4)),
  as.character(round(cor(mask$mask_rare, mask$coronavirus_virus, use = "complete.obs"),4)),
  as.character(round(cor(mask$mask_rare, mask$cloth_face_mask, use = "complete.obs"), 4)), 
  as.character(round(cor(mask$mask_rare, mask$mask, use = "complete.obs"), 4)), 
  as.character(round(cor(mask$mask_rare, mask$civil_and_political_rights, use = "complete.obs"), 4)), 

  
  "Covid Rates",NA,  NA, 'covid_19',  'smell_loss', 'taste_loss', NA, NA, 
  "Covid Rates","Covid Rate", "Pearson's R Correlation", 
  as.character(round(cor(covid$covid_rate, covid$covid_19, use = "complete.obs"),4)),
  as.character(round(cor(covid$covid_rate, covid$smell_loss, use = "complete.obs"),4)),
  as.character(round(cor(covid$covid_rate, covid$taste_loss, use = "complete.obs"),4)),
  NA,NA, 
  "Covid Rates","Covid Rate", "repeated measures correlation coefficient", 
  as.character(covid_corrs[[2,2]]), 
  as.character(covid_corrs[[2,3]]), 
  as.character(covid_corrs[[2,4]]), NA,NA, 
  
  "Suicide Rates",NA,  NA, 'suicide', 'depression', 'suicide_hotline', NA,NA,  
  "Suicide Rates","Suicide Rate", "Pearson's R Correlation",
  as.character(round(cor(suicide$death_rate, suicide$suicide, use = "complete.obs"),4)),
  as.character(round(cor(suicide$death_rate, suicide$depression, use = "complete.obs"),4)),
  as.character(round(cor(suicide$death_rate, suicide$suicide_hotline, use = "complete.obs"),4)),
  NA,NA, 
  "Suicide Rates","Suicide Rate", "repeated measures correlation coefficient", 
  as.character(suicide_corrs[[2,2]]), 
  as.character(suicide_corrs[[2,3]]), 
  as.character(suicide_corrs[[2,4]]), NA,NA, 
  
  "2016 Presidential Votes",NA,  NA, 'Hilary Clinton',  'Donald Trump', NA, NA,NA, 
  "2016 Presidential Votes","2016 Votes for Clinton", "Pearson's R Correlation",
  as.character(round(cor(pres_2016$clinton_p, pres_2016$hillary_clinton_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2016$clinton_p, pres_2016$donald_trump_trend, use = "complete.obs"),4)), NA, NA,NA, 
  "2016 Presidential Votes","2016 Votes for Trump", "Pearson's R Correlation",
  as.character(round(cor(pres_2016$trump_p, pres_2016$hillary_clinton_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2016$trump_p, pres_2016$donald_trump_trend, use = "complete.obs"),4)), NA, NA,NA, 
  
  "2020 Presidential Votes",NA,  NA, 'Joe Biden',  'Donald Trump', NA, NA,NA, 
  "2020 Presidential Votes","2020 Votes for Biden", "Pearson's R Correlation",
  as.character(round(cor(pres_2020$biden_p, pres_2020$joe_biden_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2020$biden_p, pres_2020$donald_trump_trend, use = "complete.obs"),4)), NA, NA,NA, 
  "2020 Presidential Votes","2020 Votes for Trump", "Pearson's R Correlation",
  as.character(round(cor(pres_2020$trump_p, pres_2020$joe_biden_trend, use = "complete.obs"),4)),
  as.character(round(cor(pres_2020$trump_p, pres_2020$donald_trump_trend, use = "complete.obs"),4)), NA, NA, NA 
)  %>%   
  select(grp, measure, variable, trend1, trend2, trend3, trend4, trend5) %>% 
  as_grouped_data(groups = c("grp"), columns = NULL)

as_flextable(table2, hide_grouplabel = TRUE) %>% 
  set_caption(caption = "Correlation Results")  %>% 
  # theming
  theme_box() %>% 
  bg(bg = my_palette[8], part = "header") %>% 
  color(color = "white", part = "header") %>%
  set_table_properties(layout = "autofit", width = 1) %>% 
  # grouped row style
  fontsize(size = 8) %>% 
  bold(i = ~ !is.na(grp), bold = TRUE) %>% 
  italic(i = ~ !is.na(grp), italic = TRUE) %>% 
  bg(i = ~ !is.na(grp), bg = "#bad7db", part = "body")
```


```{r vacc_hes_plot}

vacc_hes %>% 
  select(vacc_hesitant, covid_19_vaccine,
         vaccine, coronavirus_disease,
         coronavirus_virus, fips) %>% 
  pivot_longer(cols = -c(vacc_hesitant, fips)) %>% 
  ggplot(aes(x=value, y = vacc_hesitant)) + 
  geom_point(aes(group = name, color = name ), show.legend = FALSE, alpha = 0.6) +
  geom_smooth( aes(group = name),method = 'lm', fill = my_palette[10], color = my_palette[10]) +
  labs(x = '', 
       y = "Percent of adults who are vaccine hesitant")+ 
  scale_y_continuous(labels = scales::label_percent()) +
  scale_color_manual(values = my_palette[c(3,6,8,4)]) + 
  facet_wrap(~name)
```

```{r vacc_hes_analysis}
#|  results='asis'


vacc_hes <- vacc_hes %>% 
  select(-c(DMA_google, Geo_QName)) %>% 
  mutate(across(covid_19_vaccine:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

vacc_hes_1  <- lm(vacc_hesitant  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband,
                  data = vacc_hes)
vacc_hes_2 <- lm(vacc_hesitant  ~  covid_19_vaccine + vaccine + 
                   coronavirus_disease + coronavirus_virus,
                 data = vacc_hes)
vacc_hes_3  <- lm(vacc_hesitant  ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + broadband + 
                    covid_19_vaccine + vaccine + 
                   coronavirus_disease + coronavirus_virus,
                   data = vacc_hes)

modelsummary(list(vacc_hes_2, vacc_hes_1, vacc_hes_3), 
             title = "Linear Regression Results for Vaccine Hesitancy",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE'
             ) 


```



```{r mask_plot}
mask %>% 
  select(mask_rare,  coronavirus_disease,
               coronavirus_virus, cloth_face_mask, 
               mask, fips) %>% 
  pivot_longer(cols = -c(mask_rare, fips)) %>% 
  ggplot(aes(x=value, y = mask_rare, )) + 
  geom_point(aes(group = name, color = name ), show.legend = FALSE, alpha = 0.6) +
  geom_smooth( aes(group = name), method = 'lm', fill = my_palette[10], color = my_palette[10]) +
  labs(x = '', 
       y = "Percent of adults who rarely use masks")+ 
  scale_y_continuous(labels = scales::label_percent()) +
  scale_color_manual(values = my_palette[c(3,6,8,4)]) + 
  facet_wrap(~name)
```

```{r mask_analysis}
#|  results='asis'


mask <- mask %>% 
  select(-c(DMA_google, Geo_QName)) %>% 
  mutate(across(coronavirus_disease:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))


mask_1  <- lm(mask_rare  ~ 
                total_pop + pop_density + unemployment_rate + 
                over_65 + poverty_rate + median_income + broadband,
              data = mask)

mask_2 <- lm(mask_rare  ~ coronavirus_disease +
               coronavirus_virus + cloth_face_mask + 
               mask,
             data = mask)

mask_3  <- lm(mask_rare  ~ 
                total_pop + pop_density + unemployment_rate + 
                over_65 + poverty_rate + median_income + broadband + 
                coronavirus_disease +
                coronavirus_virus + cloth_face_mask + 
                mask,
              data = mask)

modelsummary(list(mask_2, mask_1, mask_3), 
             title = "Linear Regression Results for Rare Mask Usage",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE') 

```


```{r covid_plot}
covid   %>% 
    filter(date %in% lubridate::ymd(c('2020-09-14'))) %>% 
  select(covid_rate, covid_19, smell_loss, taste_loss, fips) %>% 
  pivot_longer(cols = -c(covid_rate, fips)) %>% 
  ggplot(aes(x=covid_rate, y = value, group = fips, color = name)) + 
  geom_point(aes(group = name, color = name ), show.legend = FALSE, alpha = 0.6) +
  geom_smooth( aes(group = name), method = 'lm', fill = my_palette[10], color = my_palette[10]) +
  labs(x = '', 
       y = "COVID-19 case rate on 09-14-2020")+ 
  scale_color_manual(values = my_palette[c(3,6,8,4)]) + 
  facet_wrap(~name)
```

```{r covid_analysis}
#|  results='asis'


covid <- covid %>% 
  ungroup() %>% 
  select(-c(dma_google, Geo_QName, county, state)) %>% 
  mutate(across(covid_19:broadband  , ~ scale(.x, center=TRUE, scale=TRUE)[,1])) %>% 
  left_join(covid %>%
              group_by(fips) %>%
              summarize(covid_rate_fips_mean = mean(covid_rate, na.rm = TRUE)),
            by = "fips") %>% 
  drop_na(covid_rate)


covid_1  <- lme(covid_rate  ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income + broadband +
                  covid_rate_fips_mean + date,
                random = ~ 1 | fips, na.action=na.omit,data = covid)
covid_2 <- lme(covid_rate  ~  
                 covid_19 + smell_loss + taste_loss + covid_rate_fips_mean  + 
                 date,
               random = ~ 1 | fips, na.action=na.omit, data = covid)
covid_3  <- lme(covid_rate  ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income + broadband + 
                  covid_19 + smell_loss + taste_loss + covid_rate_fips_mean  +
                  date,
                random = ~ 1 | fips, na.action=na.omit, data = covid)

modelsummary(list(covid_2, covid_1, covid_3), 
             title = "Hierarchical Model for Covid Case Rates",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Random intercept per county'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE|SD') 

# TODO Drop SD intercept and observations,

```

```{r suicide_plot}
suicide   %>% 
    filter(year == 2019) %>% 
  select(death_rate, suicide, depression, suicide_hotline, fips) %>% 
  pivot_longer(cols = -c(death_rate, fips)) %>% 
  ggplot(aes(x=death_rate, y = value, group = fips, color = name)) + 
  geom_point(aes(group = name, color = name ), show.legend = FALSE, alpha = 0.6) +
  geom_smooth( aes(group = name), method = 'lm', fill = my_palette[10], color = my_palette[10]) +
  labs(x = '', 
       y = "Death rate by suicide for 2019")+ 
  scale_color_manual(values = my_palette[c(3,6,8,4)]) + 
  facet_wrap(~name)
```

```{r suicide_analysis}
#|  results='asis'

suicide <- suicide %>% 
  select(-c(DMA_google, Geo_QName, county, population, deaths)) %>%
  ungroup() %>% 
  mutate(across(death_rate:broadband, ~ scale(.x, center=TRUE, scale=TRUE)[,1])) 

suicide <- suicide %>% 
  left_join(suicide %>%
              group_by(fips) %>%
              summarize(death_rate_fips_mean = mean(death_rate , na.rm = TRUE)),
            by = "fips") %>% 
  drop_na(death_rate )


suicide_1  <- lme(death_rate   ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income +
                  # death_rate_fips_mean +
                    year,
                random = ~ 1 | fips, na.action=na.omit,data = suicide)
suicide_2 <- lme(death_rate   ~  
                 suicide + depression + suicide_hotline + 
                   # death_rate_fips_mean  +
                 year,
               random = ~ 1 | fips, na.action=na.omit, data = suicide)
suicide_3  <- lme(death_rate   ~ 
                  total_pop + pop_density + unemployment_rate + 
                  over_65 + poverty_rate + median_income +
                  suicide + depression + suicide_hotline +
                    # death_rate_fips_mean  +
                  year,
                random = ~ 1 | fips, na.action=na.omit, data = suicide)


modelsummary(list(suicide_2, suicide_1, suicide_3), 
             title = "Hierarchical Model for Suicide Rates",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Random intercept per county'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE') 

# TODO Drop SD intercept and observations,
```

```{r pres_plot}
pres_2016_plot <- pres_2016 %>% 
  select(clinton_p, 
         # trump_p,
         donald_trump_trend, hillary_clinton_trend,
         fips) %>% 
  pivot_longer(cols = -c(clinton_p, fips), names_to = "trend", values_to = "trend_value") %>% 
  # pivot_longer(cols = -c(trend, trend_value, fips), names_to = "pres", values_to = "pres_percent") %>% 

  ggplot(aes(x=trend_value, y = clinton_p, color = trend )) + 
  geom_jitter(show.legend = FALSE, alpha = 0.6) +
  geom_smooth(method = 'lm', fill = my_palette[10], color = my_palette[10]) +
  labs(x = '', 
       y = "% of votes for Hilary Clinton, 2016")+ 
  scale_y_continuous(labels = scales::label_percent()) +
  scale_color_manual(values = my_palette[c(3,6,8,4)]) +
  facet_wrap(~ trend, scales = 'free_x')
  # facet_grid(rows = vars(pres), cols = vars(trend), scales = 'free_x')

pres_2020_plot <- pres_2020 %>% 
  select(biden_p, 
         donald_trump_trend, joe_biden_trend,
         fips) %>% 
  pivot_longer(cols = -c(biden_p, fips), names_to = "trend", values_to = "trend_value") %>% 

  ggplot(aes(x=trend_value, y = biden_p, color = trend )) + 
  geom_jitter(show.legend = FALSE, alpha = 0.6) +
  geom_smooth(method = 'lm', fill = my_palette[10], color = my_palette[10]) +
  labs(x = '', 
       y = "% of votes for Joe Biden, 2020")+ 
  scale_y_continuous(labels = scales::label_percent()) +
  scale_color_manual(values = my_palette[c(8,4)]) +
  facet_wrap(~ trend, scales = 'free_x')
  # facet_grid(rows = vars(pres), cols = vars(trend), scales = 'free_x')

library(patchwork)
pres_2016_plot /pres_2020_plot

```

```{r pres_2016_analysis}

#|  results='asis'

pres_2016 <- pres_2016 %>% 
  select(-c(Geo_QName, dma)) %>% 
  mutate(across(c(hillary_clinton_trend,donald_trump_trend:broadband)  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

vacc_hes_1a  <- lm(clinton_p   ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income,
                  data = pres_2016)
vacc_hes_2a <- lm(clinton_p   ~  
                    # donald_trump_trend + 
                    hillary_clinton_trend,
                 data = pres_2016)

vacc_hes_3a  <- lm(clinton_p   ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + 
                     # donald_trump_trend + 
                     hillary_clinton_trend,
                   data = pres_2016)

vacc_hes_1b  <- lm(trump_p    ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income,
                  data = pres_2016)
vacc_hes_2b <- lm(trump_p    ~  donald_trump_trend + hillary_clinton_trend,
                 data = pres_2016)

vacc_hes_3b  <- lm(trump_p    ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + 
                     donald_trump_trend + hillary_clinton_trend,
                   data = pres_2016)

modelsummary(list(vacc_hes_2a,vacc_hes_1a, vacc_hes_3a 
                  # ,vacc_hes_1b, vacc_hes_2b,vacc_hes_3b
                  ), 
             title = "Linear Regression Results for 2016 Presidential Election Results (Hilary Clinton Shown)",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Results predicting Donald J. Trump percentage largely equivalent and available upon request.'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE|F'
             ) 


```

```{r pres_2020_analysis}
#|  results='asis'


pres_2020 <- pres_2020 %>% 
  select(-c(Geo_QName, dma)) %>% 
  mutate(across(c(joe_biden_trend,donald_trump_trend:broadband)  , ~ scale(.x, center=TRUE, scale=TRUE)[,1]))

vacc_hes_1a  <- lm(biden_p   ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income,
                  data = pres_2020)
vacc_hes_2a <- lm(biden_p   ~  
                    # donald_trump_trend +
                    joe_biden_trend,
                 data = pres_2020)

vacc_hes_3a  <- lm(biden_p   ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + 
                     # donald_trump_trend + 
                     joe_biden_trend,
                   data = pres_2020)

vacc_hes_1b  <- lm(trump_p    ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income,
                  data = pres_2020)
vacc_hes_2b <- lm(trump_p    ~  donald_trump_trend ,
                  # +                 joe_biden_trend,
                 data = pres_2020)

vacc_hes_3b  <- lm(trump_p    ~ 
                    total_pop + pop_density + unemployment_rate + 
                    over_65 + poverty_rate + median_income + 
                     donald_trump_trend ,
                   # + joe_biden_trend,
                   data = pres_2020)

modelsummary(list(vacc_hes_2a, vacc_hes_1a, vacc_hes_3a 
                  # ,vacc_hes_1b, vacc_hes_2b,vacc_hes_3b
                  ), 
             title = "Linear Regression Results for 2020 Presidential Election Results (Joe Biden Shown)",
             notes= c('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Results predicting Donald J. Trump percentage largely equivalent and available upon request.'),
             estimate = "{estimate}{stars}",
             gof_omit = 'AIC|BIC|ICC|RMSE|F'
             ) 


```